Liveness-Based Garbage Collection for Lazy Languages

### 2015-07-10 paper/paper.tex
CFG->NFA->DFA fig fixed

\begin{figure}[t!]
  \begin{tabular}{cc}
\psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
\begin{pspicture}(0,-5)(50,12)
  \putnode{t0}{origin}{5}{0}{\var{\Df{\length}{1}}}
  \putnode{t1}{t0}{15}{0}{\pscirclebox{\mbox{\ \ \ \ }}} \hspace{5mm}
  \putnode{t2}{t1}{15}{0}{\pscirclebox[doubleline=true]{\mbox{\ \ \ \ }}}
  \psset{arrows=->}
  \ncline{t0}{t1}
  \ncline{t1}{t2}
  \putnode{l0}{t1}{7}{2}{\clazy}
  \nccurve[angleA=45, angleB=135, ncurv=4, nodesep=-1]{t1}{t1}
  \putnode{l1}{t1}{0}{10}{\acdr, \clazy}
  \nccurve[angleA=45, angleB=135, ncurv=3, nodesep=-1]{t2}{t2}
  \putnode{l2}{t2}{0}{10}{\clazy}
\end{pspicture}
&
\psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
\begin{pspicture}(0,-5)(30,12)
  \putnode{t0}{origin}{5}{0}{\var{\Df{\length}{1}}}
  \putnode{t1}{t0}{15}{0}{\pscirclebox[doubleline=true]{\mbox{\ \ \ \ }}}
  \psset{arrows=->}
  \ncline{t0}{t1}
  \nccurve[angleA=45, angleB=135, ncurv=3, nodesep=-1]{t1}{t1}
  \putnode{l1}{t1}{0}{10}{\acdr}
\end{pspicture}
\\
(a) & (b) \\
\psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
\begin{pspicture}(0,-5)(50,12) %\psframe(0,-5)(40,12)
  \putnode{t0}{origin}{5}{0}{\var{\Lanv{9}{\pa}}}
  \putnode{t1}{t0}{9}{0}{\pscirclebox{\mbox{\ \ \ \ }}} \hspace{5mm}
  \putnode{t2}{t1}{12}{0}{\pscirclebox{\mbox{\ \ \ \ }}}
  \putnode{t3}{t2}{12}{0}{\pscirclebox[doubleline=true]{\mbox{\ \ \ \ }}}
  \psset{arrows=->}
  \ncline{t0}{t1}
  \ncline{t1}{t2}
  \putnode{l0}{t1}{5}{2}{\clazy}
  \ncline{t2}{t3}
  \putnode{l1}{t2}{5}{2}{\bcar}
  \nccurve[angleA=55, angleB=125, ncurv=3, nodesep=-.8]{t3}{t3}
  \putnode{l2}{t3}{0}{10}{\acdr}
\end{pspicture}
&
\psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
\begin{pspicture}(0,-5)(20,12) %\psframe(0,-5)(40,12)
  \putnode{t0}{origin}{5}{0}{\var{\Lanv{9}{\pa}}}
  \putnode{t1}{t0}{9}{0}{\pscirclebox{\mbox{\ \ \ \ }}} \hspace{5mm}
  \psset{arrows=->}
  \ncline{t0}{t1}
\end{pspicture}
\\
(c) & (d)  \end{tabular}
\caption{(a) The  grammar rules for \var{\Df{\length}{1}}
  converted into an automaton, and simplified in (b). (c) and (d) show
  the   same  for   \var{\Lanv{9}{\pa}}.   }\label{fig:example-automata}
\figrule
\end{figure}

Figure~\ref{fig:example-automata}(a)  shows the  NFA that  is obtained
from the  grammar for  \var{\Df{\length}{1'}}.  The  NFA is
simplified by replacing the edges labeled \clazy\ by $\epsilon$. While
checking for liveness during GC, a forward access-path traversing such
an edge  will be accepted  only if it  can reach a final  state.  This
effectively implements  the $\hookrightarrow$ simplification  rule for
\clazy\ for forward paths. The resulting automaton, made deterministic,
is shown in Figure~\ref{fig:example-automata}(b). This expectedly says
that  given  a demand  $\epsilon$  on  \length,  the liveness  of  its
argument             is             $\acdr^{*}$.            Similarly,
Figure~\ref{fig:example-automata}(c)    shows   the    automaton   for
\var{\Lanv{9}{\pa}}. The simplified automaton, shown in
Figure~\ref{fig:example-automata}(d) does not accept any forward
paths. This reflects the lazy nature of our language. \length\  does not
evaluate its elements and thus \pa\ is dead at $\pi_9$.

>>>>

\begin{figure}[t!]
  \begin{tabular}{cc}
\psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
\begin{pspicture}(0,-5)(50,12)
  \putnode{t0}{origin}{5}{0}{\var{\Df{\length}{1}}}
  \putnode{t1}{t0}{15}{0}{\pscirclebox{\mbox{\ \ \ \ }}} \hspace{5mm}
  \putnode{t2}{t1}{15}{0}{\pscirclebox[doubleline=true]{\mbox{\ \ \ \ }}}
  \psset{arrows=->}
  \ncline{t0}{t1}
  \ncline{t1}{t2}
  \putnode{l0}{t1}{7}{2}{\clazy}
  \nccurve[angleA=45, angleB=135, ncurv=4, nodesep=-1]{t1}{t1}
  \putnode{l1}{t1}{0}{8}{\acdr, \clazy}
  \nccurve[angleA=45, angleB=135, ncurv=3, nodesep=-1]{t2}{t2}
  \putnode{l2}{t2}{0}{8}{\clazy}
\end{pspicture}
&
\psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
\begin{pspicture}(0,-5)(30,12)
  \putnode{t0}{origin}{5}{0}{\var{\Df{\length}{1}}}
  \putnode{t1}{t0}{15}{0}{\pscirclebox[doubleline=true]{\mbox{\ \ \ \ }}}
  \psset{arrows=->}
  \ncline{t0}{t1}
  \nccurve[angleA=45, angleB=135, ncurv=3, nodesep=-1]{t1}{t1}
  \putnode{l1}{t1}{0}{8}{\acdr}
\end{pspicture}
\\
(a) & (b) \\
\psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
\begin{pspicture}(0,-5)(50,12) %\psframe(0,-5)(40,12)
  \putnode{t0}{origin}{5}{0}{\var{\Lanv{9}{\pa}}}
  \putnode{t1}{t0}{9}{0}{\pscirclebox{\mbox{\ \ \ \ }}} \hspace{5mm}
  \putnode{t2}{t1}{12}{0}{\pscirclebox{\mbox{\ \ \ \ }}}
%  \putnode{t3}{t2}{12}{0}{\pscirclebox[doubleline=true]{\mbox{\ \ \ \ }}}
  \psset{arrows=->}
  \ncline{t0}{t1}
  \ncline{t1}{t2}
  \putnode{l0}{t1}{5}{2}{\clazy}
%  \ncline{t2}{t3}
%  \nccurve[angleA=55, angleB=125, ncurv=3, nodesep=-.8]{t3}{t3}
%  \putnode{l2}{tD1}{0}{10}{\acdr}
 \putnode{tD1}{t2}{12}{0}{\pscirclebox{\mbox{\ \ \ \ }}} \hspace{5mm}
  \putnode{tD2}{tD1}{12}{0}{\pscirclebox[doubleline=true]{\mbox{\ \ \ \ }}}
  \psset{arrows=->}
  \ncline{t2}{tD1}
  \putnode{l1}{t2}{5}{2}{\bcar}
  \ncline{tD1}{tD2}
  \putnode{lD0}{tD1}{5}{2}{\clazy}
  \nccurve[angleA=45, angleB=135, ncurv=4, nodesep=-1]{tD1}{tD1}
  \putnode{lD1}{tD1}{0}{8}{\acdr, \clazy}
  \nccurve[angleA=45, angleB=135, ncurv=3, nodesep=-1]{tD2}{tD2}
  \putnode{lD2}{tD2}{0}{8}{\clazy}
\end{pspicture}
&
\psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
\begin{pspicture}(0,-5)(20,12) %\psframe(0,-5)(40,12)
  \putnode{t0}{origin}{5}{0}{\var{\Lanv{9}{\pa}}}
  \putnode{t1}{t0}{9}{0}{\pscirclebox{\mbox{\ \ \ \ }}} \hspace{5mm}
  \psset{arrows=->}
  \ncline{t0}{t1}
\end{pspicture}
\\
(c) & (d)  \end{tabular}
\caption{(a) The  grammar rules for \var{\Df{\length}{1}}
  converted into an automaton, and simplified in (b). The same for \var{\Lanv{9}{\pa}}
  is shown in (c) and (d)}\label{fig:example-automata}
\figrule
\end{figure}

Figure~\ref{fig:example-automata}(a)  shows the  NFA that  is obtained
from the  grammar for  \var{\Df{\length}{1}}.  The  NFA is
simplified by replacing the edges labeled \clazy\ by $\epsilon$. While
checking for liveness during GC, a forward access-path traversing such
an edge  will be accepted  only if it  can reach a final  state.  This
effectively implements  the $\hookrightarrow$ simplification  rule for
\clazy\ for forward paths. The resulting automaton, made deterministic,
is shown in Figure~\ref{fig:example-automata}(b). This expectedly says
that  given  a demand  $\epsilon$  on  \length,  the liveness  of  its
argument             is             $\acdr^{*}$.            Similarly,
Figure~\ref{fig:example-automata}(c)    shows   the    automaton   for
\var{\Lanv{9}{\pa}}. The simplified automaton, shown in
Figure~\ref{fig:example-automata}(d) does not accept any forward
paths. This reflects the lazy nature of our language. \length\  does not
evaluate its elements and thus \pa\ is dead at $\pi_9$.
